### 2024-01-31 outputs/paper/paper.qmd
setup with example starter file

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-02-11 outputs/paper/paper.qmd
introduction updated

@@
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Ziheng Zhong
+  - Charles Cai
+thanks: "Code and data are available at: https://github.com/iJustinn/Effect_of_Charter_Schools_on_School_Segregation.git"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Write after we have the results"
 format: pdf
@@
 
-You can and should cross-reference sections and sub-sections.
+In recent years, charter schools have become a popular choice for families looking for alternatives to traditional public schools in the United States.
+These schools are praised by many for offering new opportunities for education improvement, suggesting that they can lead to better student performance by encouraging competition among schools.
+Research has shown that charter schools can indeed help students achieve higher test scores.
+However, not all parents choose schools based on their effectiveness alone, which can lead to increased differences in school quality and potentially make educational inequality worse.
+
+An important issue in American education is the ongoing problem of racial segregation in schools, a problem that persists over 65 years since the historic Brown v.
+Board of Education decision aimed to end it.
+The question then arises: how do charter schools, which allow families to choose their schools, affect this racial divide?
+The answer is not straightforward.
+On one hand, charter schools could help mix students from different backgrounds by not tying school attendance to where a family lives, which often reflects racial divisions.
+On the other hand, if families select schools based on racial preferences or other factors, charter schools could actually make segregation worse.
+
+While some charter schools aim to serve underrepresented communities, the overall impact of charter schools on racial segregation in public schools has been unclear and somewhat contested.
+Previous research on this topic has been mixed, with some studies suggesting charter schools might concentrate minority students more than traditional public schools do, while others find little difference in racial isolation between charter and traditional schools.
+
+This paper aims to shed light on this issue by providing detailed analysis on whether charter schools contribute to or alleviate racial segregation in public schools.
+We use a large dataset covering nearly all public schools in the country from 1998 to 2018 to examine how the presence of charter schools affects the racial makeup of schools.
+Our study looks at how these effects play out across different regions and for different racial and ethnic groups, using sophisticated methods to ensure our findings are reliable.
+
+We find that the expansion of charter schools has, on average, led to more segregation among racial and ethnic groups in schools, particularly in urban areas with large Black and Hispanic populations.
+However, the situation is complex, as charter schools also seem to bring some level of integration across larger geographic areas like metropolitan regions.
+
+Our research contributes to the ongoing debate on the role of charter schools in American education, highlighting the need for careful consideration of how school choice impacts racial segregation.
+While charter schools have the potential to offer important benefits, it's crucial to understand and address the unintended consequences they may have on racial dynamics within the education system.
 
@@
 
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
 # Model
-
-The goal of our modelling strategy is twofold. Firstly,...
 
@@
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
 ### Model justification
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
 # Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
 

### 2024-02-11 outputs/paper/paper.qmd
paper updated

@@
 format: pdf
+toc: true
+toc-depth: 3
 number-sections: true
@@
 library(tidyverse)
-library(palmerpenguins)
+library(ggplot2)
+library(janitor)
+library(dplyr)
+library(readr)
+library(knitr)
 ```
@@
 
-# Results
+# Results {#sec-res}
 
-# Discussion
+# Discussion {#sec-dis}
 

### 2024-02-12 outputs/paper/paper.qmd
introduction updated

@@
   - Charles Cai
-thanks: "Code and data are available at: https://github.com/iJustinn/Effect_of_Charter_Schools_on_School_Segregation.git"
+thanks: "Code and data are available at: https://github.com/iJustinn/Impact_of_Past_Behavior_Normality_on_Regret.git"
 date: today
@@
 
-In recent years, charter schools have become a popular choice for families looking for alternatives to traditional public schools in the United States.
-These schools are praised by many for offering new opportunities for education improvement, suggesting that they can lead to better student performance by encouraging competition among schools.
-Research has shown that charter schools can indeed help students achieve higher test scores.
-However, not all parents choose schools based on their effectiveness alone, which can lead to increased differences in school quality and potentially make educational inequality worse.
+This study delves into how people make decisions, feel about their choices, and, especially, how they deal with regret when things don't go as planned.
+Imagine taking a different route home and getting into a car accident.
+How much more would you regret these decisions just because they're not what you usually do?
+This idea stems from a concept called norm theory, which suggests that people feel extra bad about outcomes when they result from actions that are not typical for them.
 
-An important issue in American education is the ongoing problem of racial segregation in schools, a problem that persists over 65 years since the historic Brown v.
-Board of Education decision aimed to end it.
-The question then arises: how do charter schools, which allow families to choose their schools, affect this racial divide?
-The answer is not straightforward.
-On one hand, charter schools could help mix students from different backgrounds by not tying school attendance to where a family lives, which often reflects racial divisions.
-On the other hand, if families select schools based on racial preferences or other factors, charter schools could actually make segregation worse.
+This paper focuses on two main experiments to unravel this complex bundle of decisions and emotions.
+The first experiment places people in a scenario where they decide whether or not to pick up a hitchhiker, testing how much one would regret the decision based on whether it's something they normally do.
+The second experiment changes the setting to driving, where choosing a different route leads to an accident, and looks at how this break from routine affects their feelings of regret.
 
-While some charter schools aim to serve underrepresented communities, the overall impact of charter schools on racial segregation in public schools has been unclear and somewhat contested.
-Previous research on this topic has been mixed, with some studies suggesting charter schools might concentrate minority students more than traditional public schools do, while others find little difference in racial isolation between charter and traditional schools.
+At the center of the investigation is the idea of exceptionality—making choices that break from people's usual patterns—and how this leads to stronger feelings of regret when things go wrong.
+This paper also explores whether people think that making an unusual choice (one they had full control over) makes them more responsible for the bad outcome, compared to if someone else made the decision for them or if it was just another day doing the same old thing.
 
-This paper aims to shed light on this issue by providing detailed analysis on whether charter schools contribute to or alleviate racial segregation in public schools.
-We use a large dataset covering nearly all public schools in the country from 1998 to 2018 to examine how the presence of charter schools affects the racial makeup of schools.
-Our study looks at how these effects play out across different regions and for different racial and ethnic groups, using sophisticated methods to ensure our findings are reliable.
+This paper takes a closer look at the concept of free will in people's choices, whether they believe that they had the freedom to choose differently in situations they set up themselves versus those set up by others or by habit.
+This part of the study helps people understand better why one might blame themselves more when they voluntarily make a choice that leads to regret.
 
-We find that the expansion of charter schools has, on average, led to more segregation among racial and ethnic groups in schools, particularly in urban areas with large Black and Hispanic populations.
-However, the situation is complex, as charter schools also seem to bring some level of integration across larger geographic areas like metropolitan regions.
-
-Our research contributes to the ongoing debate on the role of charter schools in American education, highlighting the need for careful consideration of how school choice impacts racial segregation.
-While charter schools have the potential to offer important benefits, it's crucial to understand and address the unintended consequences they may have on racial dynamics within the education system.
+By picking apart these scenarios, this paper aims to shed light on the tricky relationship between stepping out of one's comfort zone in decision-making and the regret that can follow, especially when things don't turn out well.
+The goal of this research is to contribute to the broader conversation on how people's minds work in these situations, improving understanding of why people feel the way they do after making certain choices.
+This insight is not just for psychologists but for anyone interested in the little decisions people make every day and how they shape their feelings and lives.
 

### 2024-02-13 outputs/paper/Paper_past.qmd
introduction updated, data source updated

---
title: "Continuous Rise in Crime Cases Reported in Toronto Over Recent Years"
subtitle: "An Analysis of Toronto's Reported Crime from 2014 to 2022"
author: 
  - Ziheng Zhong
thanks: "Code and data are available at: https://github.com/iJustinn/Totonto_Reported_Crimes.git"
date: today
date-format: long
abstract: "This study presents a comprehensive analysis of reported crime data in Toronto, as sourced from Open Data Toronto, to explore the trend of urban criminal patterns in Toronto over the past years. By examining reported incidents from 2014 to 2022, this paper identifies key trends in crime cases across different years and police divisions, highlighting possible key factors that may caused such results. In a word, the paper seeks to provide actionable insights for city planners, law enforcement, and community advocates, aiming to build a safer, more enjoyable urban living environment."
format: pdf
toc: true
toc-depth: 3
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(opendatatoronto)
library(tidyverse)
library(ggplot2)
library(janitor)
library(dplyr)
library(readr)
library(knitr)

# read the cleaned_data
raw_data <- read_csv(here::here("inputs/data/raw_data.csv"))

# read the cleaned_data
cleaned_data <- read_csv(here::here("outputs/data/cleaned_data.csv"))

# read the sum_by_year_data
sum_by_year_data <- read_csv(here::here("outputs/data/sum_by_year_data.csv"))

# read the sum_by_division_data
sum_by_division_data <- read_csv(here::here("outputs/data/sum_by_division_data.csv"))
```

# Introduction

Public safety, crucial for ensuring high-quality life, requires continuous evaluation and enforcement. The city of Toronto, with one of Canada's largest populations, is not immune to the challenges posed by crime. This paper takes a deep dive into the extensive reported crime data provided by Open Data Toronto[@citeODT], covering the period from 2014 to 2022. The analysis is grounded in a methodical approach to identify and interpret crime trends across the past years, also within the city's diverse police divisions, offering a detailed chronicle of criminal activity over a selected time span. More specifically, this research analyzed and visualized how crime is distributed across different divisions and how it changes over the years. By dissecting the data year by year, division by division, this study uncovers patterns that could inform future urban planning and law enforcement strategies. The study considers the role of population growth, natural diseases, and urban development as potential factors influencing these patterns.

Moreover, the paper aims to go beyond merely displaying statistics. It endeavors to contextualize the numerical findings within the broader spectrum of Toronto's urban structure. This discussion is motivated by the observation that changes in crime rates often correlate with shifts in other city indicators, such as education level[@citeEdu], unemployment rate[@citeUnemp], economic level[@citeEcon], and so on. By doing this, the study aims to offer useful advice that can help city officials and the police create plans to make the city safer and better for all residents. In this manner, the paper serves not merely as an exposition of data but as a roadmap towards a better living environment.

The subsequent sections of this paper are organized to facilitate a comprehensive understanding of the study and its implications. After this introduction, @sec-2 outlines the methodology used in the analysis, emphasizing transparency and replicability. @sec-3 presents the findings in detail, charting the trends in reported crimes over time and across Toronto's police divisions. @sec-4 discusses these findings from various perspectives, including natural diseases, infrastructure construction, and the city's structure, among others. Finally, @sec-5 concludes the paper, summarizing the key insights and findings of this research.

# Data {#sec-2}

## Data Source

The data used in this paper was collected by the OpenDataToronto Library [@citeODT]. The specific data set used in this research is the 'police-annual-statistical-report-reported-crimes' [@citeDataset].

Data used in this paper was downloaded, cleaned and analyzed with the programming language R [@citeR]. Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `dplyr` [@citeDplyr], `readr` [@citeReadr], `knitr` [@citeknitr].

## Data Processing

### Data Cleaning

```{r}
#| message: false
#| echo: false
#| label: tbl-raw_data
#| tbl-cap: Head of Raw Toronto Reported Crimes Data

head (raw_data, 6) |>
  kable(
    col.names = c("ID", "Reported Year", "Division", "Category", "Sub Type", "Total Case", "Cleared Case"),
  booktabs = TRUE
  )
```

```{r}
#| message: false
#| echo: false
#| label: tbl-cleaned_data
#| tbl-cap: Head of Cleaned Toronto Reported Crimes Data

head (cleaned_data, 6) |>
  kable(
    col.names = c("Reported Year", "Division", "Total Case"),
  booktabs = TRUE
  )
```

The raw data set, first few lines shown in @tbl-raw_data, contains more than 30000 reported crime cases in Toronto from 2014 to 2022 in all police divisions, including reported crime types, case status, case ID, etc. Since this research only focusing on the division and reported year of each case, thus all other extra information in the data was cleaned as needed. Now the cleaned data only shows the number of reported crime cases in corresponding years and the police division to each case was reported. The first few lines of the cleaned data were shown in the @tbl-cleaned_data.

### Further Data Processing

In order to make it easier to draw the data in the future, the cleaned data has been further processed into data with only two variables, which are used to correspond to the x-axis and y-axis when drawing.

```{r}
#| message: false
#| echo: false
#| label: tbl-by_year_data
#| tbl-cap: Head of Toronto Reported Crimes Data by year

head (sum_by_year_data, 6) |>
  kable(
    col.names = c("Reported Year","Total Cases"),
  booktabs = TRUE
  )
```

As mentioned earlier, this paper will first analyze the trends of reported crimes in different years. Therefore, the first further processed data will add up all data from different policy divisions in each year to form a total number of reported crimes for a specific year. The head of these data was shown in @tbl-by_year_data.

```{r}
#| message: false
#| echo: false
#| label: tbl-by_division_data
#| tbl-cap: Head of Toronto Reported Crimes Data by police division

head (sum_by_division_data, 6) |>
  kable(
    col.names = c("Division","Total Cases"),
  booktabs = TRUE
  )
```

Then, this paper will analyze the total number of reported crimes received for different policy divisions disregard of the year. Therefore, the process of this data ignores the year and adds up all the data received by each policy division in the cleaned data set to form the data set shown in @tbl-by_division_data.

# Results {#sec-3}

## Across years

```{r}
#| label: fig-by_year
#| fig-cap: 2014 to 2022 Toronto Total Reported Crime Cases by Year
#| echo: false

# create the bar chart with y-axis showing total cases in the year and x-axis showing the year number
ggplot(sum_by_year_data, aes(x = report_year, y = total_cases)) +
  scale_x_continuous(breaks = sum_by_year_data$report_year) +
  geom_bar(stat = "identity", fill = "darkgrey") +
  theme_minimal() +
  labs(x = "Report Year",
       y = "Total Cases") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The bar chart (@fig-by_year) displays the total number of reported crime cases in Toronto for each year from 2014 to 2022 with the vertical axis indicates the number of cases. From a visual inspection, the following trend can be observed: starting in 2014, there appears to be a relatively steady or slight increase in the number of cases, reaching a peak around 2017 and 2018. After this peak, there is a significant decrease in the number of reported crime cases, showing almost no upward trend up to 2021.

Notably, there are two significant drops in the total reported crime cases in 2020 and 2022 compared to their previous years. These declines could be of particular interest and are worth further investigation to understand the underlying factors contributing to these changes.

## Across divisions

```{r}
#| label: fig-by_division
#| fig-cap: 2014 to 2022 Toronto Total Reported Crime Cases by Division
#| echo: false

# create the bar chart with y-axis showing total cases in the division and x-axis showing the division number
ggplot(sum_by_division_data, aes(x = division, y = total_cases, fill = division)) +
  #geom_bar(stat = "identity", position = position_dodge()) +
  geom_bar(stat = "identity", fill = "darkgrey") +
  theme_minimal() +
  labs(x = "Police Division",
       y = "Total Cases") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The bar chart (@by_division) provides a visual representation of the total reported crime cases in Toronto from 2014 to 2022, broken down by police division. Each bar corresponds to a specific division, labeled D11 through D55, along with an 'NSA' category at the end. The vertical axis quantifies the total cases.

The divisional data shows variability in the number of reported crimes, with some divisions displaying notably higher totals than others. For instance, certain divisions such as D14, D32, D51, D52, and D55, exhibit particularly tall bars, indicating a higher total of reported cases. Conversely, divisions like D12, and D13, show considerably fewer reported cases, as reflected by their shorter bars.

# Discussion {#sec-4}

## Across years

Continuing with the analysis of the chart (@fig-by_year), as mentioned in the Results section, there were two major drops in the trend line. To begin with, the drop in 2020 appears to have an obvious cause. Starting from the end of 2019[@citeCovidTimeline], the world entered a period of over two years of the pandemic. Public health measures such as lockdowns, physical distancing, and restrictions on gatherings drastically changed social behaviors and daily activities. This unprecedented situation has had a ripple effect across various sectors, impacting public safety and crime rates.

As communities went into isolation and businesses either temporarily closed or shifted to remote operations, the opportunities for certain types of crimes, such as burglaries and thefts in commercial areas, may have diminished[@citeDecreaseCrime]. Conversely, there have been concerns about a potential rise in domestic incidents and cybercrimes, as more people stayed at home and activities shifted online.

The impact of the pandemic on mental health[@citeCovidMental], driven by prolonged social isolation and economic stress, cannot be understated. These factors have influenced community dynamics and possibly crime patterns[@citeMentalCrime]. Law enforcement agencies and social services have had to adapt, sometimes shifting their focus to community support and online safety initiatives. The extended period of the pandemic has also challenged traditional approaches to crime prevention and response, prompting a reevaluation of resource allocation and policing methods.

## Across divisions

The bar chart (@fig-by_division) illustrates the total reported crime cases in Toronto by police division from 2014 to 2022. Upon examining specific divisions, certain patterns emerge. The map[@citeDivisionMap] provided by the Toronto Police Service (TPS) shows that downtown Toronto, particularly Division 51, has a significant number of reported cases. Notably, Division 51 has one of the highest counts of total crime cases on the chart. This finding strongly suggests a correlation between the high population density and activity levels in downtown areas and the rates of crime[@citeDensityCrime].

Not only in downtown Toronto but also in areas hosting large shopping malls, such as North York's Division 32, the total crime cases exceed those of other divisions significantly. Additionally, divisions characterized by high population density, such as Division 52—where the CN Tower and Scotiabank Arena are located—and Division 14, known for its numerous train stations, also report a high number of crime cases.

Conversely, Scarborough's Division 43, while reporting a considerable number of cases, does not reach the high levels observed in the aforementioned divisions. This may reflect the predominant house type (town houses rather than condos) and development situation in the area. Once again, this suggests that population density may be a factor influencing crime rates[@citeDensityCrime]. While other factors such as socioeconomic status, community engagement, and policing strategies are also significant, further information is necessary for a more comprehensive assessment.

# Conclusion {#sec-5}

"In conclusion, this paper reflects on the significant findings from the analysis of reported crime data in Toronto between 2014 and 2022. The data reveal fluctuating crime rates, with notable peaks and troughs that correspond to various factors and events, most prominently the COVID-19 pandemic. This period notably witnessed changes in crime patterns due to altered social behaviors and shifts in policing strategies. The division-specific analysis highlighted the impact of population density and the level of development on crime distribution. By leveraging detailed crime data, Toronto can continue to refine its urban safety strategies, aiming to create a city that is not only safer but also more equitable and responsive to the needs of its rapidly growing population[@citePopulation].

# Bibliography

### 2024-02-13 outputs/paper/paper.qmd
introduction updated, data source updated

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Research on The Impact of Past Behaviour Normality on Regret"
+subtitle: "The replication of two major experiments"
 author: 
   - Ziheng Zhong
-  - Charles Cai
+  - Yihang Cai
 thanks: "Code and data are available at: https://github.com/iJustinn/Impact_of_Past_Behavior_Normality_on_Regret.git"
@@
 date-format: long
-abstract: "Write after we have the results"
+abstract: "To be written after we finished the Discussion section"
 format: pdf
@@
 
-This study delves into how people make decisions, feel about their choices, and, especially, how they deal with regret when things don't go as planned.
-Imagine taking a different route home and getting into a car accident.
-How much more would you regret these decisions just because they're not what you usually do?
+Descriptive norms and injunctive norms are key concepts in understanding social behavior, each describing different aspects of how individuals perceive and are influenced by the norms within their community or society.
+Descriptive norms are about observing and following what most people do in a given situation, essentially reflecting the common actions or behaviors within a group.
+Injunctive norms, on the other hand, are about what people believe they should or shouldn't do, based on the group's values or expectations for behavior.
+These norms carry a sense of approval or disapproval, guiding behavior by suggesting what is socially acceptable or frowned upon.
+Both types of norms are powerful in shaping individual actions by providing a roadmap of 'what is done' and 'what ought to be done,' helping maintain social order and cohesion by aligning personal conduct with group expectations.
+
+This study delves into how people make decisions, feel about their choices, and, how they deal with regret when things don't go as planned.
+How much more would people regret their decisions just because they're not what you usually do?
 This idea stems from a concept called norm theory, which suggests that people feel extra bad about outcomes when they result from actions that are not typical for them.
-
-This paper focuses on two main experiments to unravel this complex bundle of decisions and emotions.
+Two main experiments was introduced to unravel this complex bundle of decisions and emotions.
 The first experiment places people in a scenario where they decide whether or not to pick up a hitchhiker, testing how much one would regret the decision based on whether it's something they normally do.
 The second experiment changes the setting to driving, where choosing a different route leads to an accident, and looks at how this break from routine affects their feelings of regret.
-
-At the center of the investigation is the idea of exceptionality—making choices that break from people's usual patterns—and how this leads to stronger feelings of regret when things go wrong.
-This paper also explores whether people think that making an unusual choice (one they had full control over) makes them more responsible for the bad outcome, compared to if someone else made the decision for them or if it was just another day doing the same old thing.
-
-This paper takes a closer look at the concept of free will in people's choices, whether they believe that they had the freedom to choose differently in situations they set up themselves versus those set up by others or by habit.
-This part of the study helps people understand better why one might blame themselves more when they voluntarily make a choice that leads to regret.
 
@@
 This insight is not just for psychologists but for anyone interested in the little decisions people make every day and how they shape their feelings and lives.
+With all been said, we predicted that someone who often picks up hitchhikers would be seen as doing something unusual (against descriptive norms) and might be judged more harshly by others (against injunctive norms).
+We also introduced a new aspect to the original experiment, looking at how these social norms and a person's usual actions stir up feelings, both good and bad.
 
-The remainder of this paper is structured as follows. @sec-data....
+The remainder of this paper is structured as follows.
+
+@sec-data
+@sec-res
+@sec-dis
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+Data used in this paper was cleaned and processed with the programming language R [@citeR].
+Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `dplyr` [@citeDplyr], `readr` [@citeReadr], `knitr` [@citeknitr].
 
-# Model
+## Source
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The data used in this paper was a replication of experimental data presented in the @citeOriginalPaper paper.
 
-## Model set-up
+## Method
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+### Experiment #1
 
-### Model justification
+### Experiment #2
+
+## Attibutes
+
+### Experiment #1
+
+### Experiment #2
 
@@
 
+## Experiment #1
+
+## Experiment #2
+
 # Discussion {#sec-dis}
 
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## First discussion point
 
@@
 ## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
 

### 2024-02-13 outputs/paper/paper.qmd
Update paper.qmd
graphs added

@@
 ```{r}
-#| include: false
-#| warning: false
 #| message: false
+#| echo: false
 
+# load packages
 library(tidyverse)
@@
 library(knitr)
+library(kableExtra)
+library(reshape2)
+```
+
+```{r}
+#| message: false
+#| echo: false
+
+# read in data
+file_path <- "../../inputs/data/raw_data.csv"
+data <- read_csv(file_path, show_col_types = FALSE)
 ```
@@
 
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: tbl-experiment1
+#| tbl-cap: Result of experiment 1
+
+# Filter the data to drop NA values
+data_filtered_1 <- data %>%
+  filter(!is.na(Sc1_regret) & !is.na(sc1_socnorms1) & !is.na(sc1_socnorms2) & !is.na(sc1_combinednorms))
+
+# Calculate the total counts and counts for Mr.Jones (1) and Mr.Smith (2)
+total_counts_1 <- nrow(data_filtered_1)
+mr_jones_counts <- sum(data_filtered_1$Sc1_regret == 1)
+mr_smith_counts <- total_counts_1 - mr_jones_counts
+
+# Calculate the counts and percentages for Mr.Jones
+mr_jones_data <- data.frame(
+  Group = "Mr.Jones(exception)",
+  Regret = paste(mr_jones_counts, "(", round((mr_jones_counts / total_counts_1) * 100, 1), "%)", sep = ""),
+  Injunctive = paste(sum(data_filtered_1$sc1_socnorms1 == 1), "(", round((sum(data_filtered_1$sc1_socnorms1 == 1) / total_counts_1) * 100, 1), "%)", sep = ""),
+  Descriptive = paste(sum(data_filtered_1$sc1_socnorms2 == 1), "(", round((sum(data_filtered_1$sc1_socnorms2 == 1) / total_counts_1) * 100, 1), "%)", sep = ""),
+  `Negative Effect` = paste(sum(data_filtered_1$sc1_combinednorms == 1), "(", round((sum(data_filtered_1$sc1_combinednorms == 1) / total_counts_1) * 100, 1), "%)", sep = "")
+)
+
+# Calculate the counts and percentages for Mr.Smith
+mr_smith_data <- data.frame(
+  Group = "Mr.Smith(routine)",
+  Regret = paste(mr_smith_counts, "(", round((mr_smith_counts / total_counts_1) * 100, 1), "%)", sep = ""),
+  Injunctive = paste(sum(data_filtered_1$sc1_socnorms1 == 2), "(", round((sum(data_filtered_1$sc1_socnorms1 == 2) / total_counts_1) * 100, 1), "%)", sep = ""),
+  Descriptive = paste(sum(data_filtered_1$sc1_socnorms2 == 2), "(", round((sum(data_filtered_1$sc1_socnorms2 == 2) / total_counts_1) * 100, 1), "%)", sep = ""),
+  `Negative Effect` = paste(sum(data_filtered_1$sc1_combinednorms == 2), "(", round((sum(data_filtered_1$sc1_combinednorms == 2) / total_counts_1) * 100, 1), "%)", sep = "")
+)
+
+# Combine the data for Mr. Jones and Mr. Smith into one dataframe
+combined_data <- rbind(mr_jones_data, mr_smith_data)
+
+# Now use kable to create the table with the combined data
+kable_output <- kable(combined_data, "latex", booktabs = TRUE) %>%
+  kable_styling(full_width = FALSE, font_size = 7)
+
+kable_output
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: fig-experiment1bar
+#| fig-cap: Bar plot of experiment 1
+
+smith_regret_p <- round((mr_smith_counts / total_counts_1) * 100, 1)
+smith_injunctive_p <- round((sum(data_filtered_1$sc1_socnorms1 == 2) / total_counts_1) * 100, 1)
+smith_descriptive_p <- round((sum(data_filtered_1$sc1_socnorms2 == 2) / total_counts_1) * 100, 1)
+smith_negative_p <- round((sum(data_filtered_1$sc1_combinednorms == 2) / total_counts_1) * 100, 1)
+jones_regret_p <- round((mr_jones_counts / total_counts_1) * 100, 1)
+jones_injunctive_p <- round((sum(data_filtered_1$sc1_socnorms1 == 1) / total_counts_1) * 100, 1)
+jones_descriptive_p <- round((sum(data_filtered_1$sc1_socnorms2 == 1) / total_counts_1) * 100, 1)
+jones_negative_p <- round((sum(data_filtered_1$sc1_combinednorms == 1) / total_counts_1) * 100, 1)
+
+# Create a data frame with the provided data
+data <- data.frame(
+  Category = c("Regret", "Injunctive Social Norm", "Descriptive Social Norm", "Negative Affect"),
+  Smith_Routine = c(smith_regret_p, smith_injunctive_p, smith_descriptive_p, smith_negative_p),
+  Jones_Exception = c(jones_regret_p, jones_injunctive_p, jones_descriptive_p, jones_negative_p)
+)
+
+# Convert the data from wide to long format
+data_long <- melt(data, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")
+
+# Reorder the factor levels of Category so that 'Regret' comes first
+data_long$Category <- factor(data_long$Category, levels = c("Regret", "Injunctive Social Norm", "Descriptive Social Norm", "Negative Affect"), 
+                             labels = c("Regret", "Social Norm\nInjunctive", 
+                                        "Social Norm\nDescriptive", "Negative\nAffect"))
+
+# Create the bar plot with the categories reordered
+ggplot(data_long, aes(x = Category, y = Percentage, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.7) +
+  scale_fill_manual(values = c("Smith_Routine" = "grey", "Jones_Exception" = "#505050"), 
+                    labels = c("Mr.Smith(Routine)", "Mr.Jones(Exception)")) +
+  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20)) +
+  theme_minimal() +
+  labs(x = "", y = "Percentage", fill = "Condition") +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, colour = "black",size = 9), 
+    legend.position = "top", legend.title = element_blank())
+```
+
 ### Experiment #2
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: tbl-experiment2
+#| tbl-cap: Result of experiment 2
+
+# Filter the data to drop NA values
+data_filtered_2 <- data %>%
+  filter(!is.na(Sc2_regret) & !is.na(Sc2_random_1) & !is.na(Sc2_random_2) & !is.na(Sc2_lucky))
+
+# Calculate the total counts and counts for Mr.Adams (1) and Mr.White (2)
+total_counts_2 <- nrow(data_filtered_2)
+mr_adams_counts <- sum(data_filtered_2$Sc2_regret == 1)
+mr_white_counts <- total_counts_2 - mr_adams_counts
+
+# Calculate the counts and percentages for Mr.Adams
+mr_adams_data <- data.frame(
+  Group = "Mr.Adams(routine)",
+  Regret = paste(mr_adams_counts, "(", round((mr_adams_counts / total_counts_2) * 100, 1), "%)", sep = ""),
+  Luck = paste(sum(data_filtered_2$Sc2_lucky == 1), "(", round((sum(data_filtered_2$Sc2_lucky == 1) / total_counts_2) * 100, 1), "%)", sep = "")
+)
+
+# Calculate the counts and percentages for Mr. White
+mr_white_data <- data.frame(
+  Group = "Mr.White(exception)",
+  Regret = paste(mr_white_counts, "(", round((mr_white_counts / total_counts_2) * 100, 1), "%)", sep = ""),
+  Luck = paste(sum(data_filtered_2$Sc2_lucky == 2), "(", round((sum(data_filtered_2$Sc2_lucky == 2) / total_counts_2) * 100, 1), "%)", sep = "")
+)
+
+# Combine the data for Mr. Jones and Mr. Smith into one dataframe
+combined_data <- rbind(mr_adams_data, mr_white_data)
+
+# Now use kable to create the table with the combined data
+kable_output <- kable(combined_data, "latex", booktabs = TRUE) %>%
+  kable_styling(full_width = FALSE, font_size = 7)
+
+kable_output
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: fig-experiment2bar
+#| fig-cap: Bar plot of experiment 2
+
+adams_regret_p <- round((mr_adams_counts / total_counts_2) * 100, 1)
+adams_luck_p <- round((sum(data_filtered_2$Sc2_lucky == 1) / total_counts_2) * 100, 1)
+white_regret_p <- round((sum(data_filtered_2$Sc2_regret == 2) / total_counts_2) * 100, 1)
+white_luck_p <- round((sum(data_filtered_2$Sc2_lucky == 2) / total_counts_2) * 100, 1)
+
+# Create a data frame with the provided data
+data <- data.frame(
+  Category = c("Regret", "Luck"),
+  Adams_Routine = c(adams_regret_p, adams_luck_p),
+  White_Exception = c(white_regret_p, white_luck_p)
+)
+
+# Convert the data from wide to long format
+data_long <- melt(data, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")
+
+# Reorder the factor levels of Category so that 'Regret' comes first
+data_long$Category <- factor(data_long$Category, levels = c("Regret", "Luck"))
+
+# Create the bar plot with the categories reordered
+ggplot(data_long, aes(x = Category, y = Percentage, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.7) +
+  scale_fill_manual(values = c("Adams_Routine" = "grey", "White_Exception" = "#505050"), 
+                    labels = c("Mr.Adams(Routine)", "Mr.White(Exception)")) +
+  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20)) + # label the y-axis every 20%
+  theme_minimal() +
+  labs(x = "", y = "Percentage", fill = "Condition") +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, colour = "black",size = 9), 
+    legend.position = "top", legend.title = element_blank())
+```
 

### 2024-02-14 outputs/paper/paper.qmd
graphs uploaded

@@
 # Create a data frame with the provided data
-data <- data.frame(
+df <- data.frame(
   Category = c("Regret", "Injunctive Social Norm", "Descriptive Social Norm", "Negative Affect"),
@@
 # Convert the data from wide to long format
-data_long <- melt(data, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")
+data_long <- melt(df, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")
 

### 2024-02-14 outputs/paper/paper.qmd
introduction modified

@@
 
-Descriptive norms and injunctive norms are key concepts in understanding social behavior, each describing different aspects of how individuals perceive and are influenced by the norms within their community or society.
-Descriptive norms are about observing and following what most people do in a given situation, essentially reflecting the common actions or behaviors within a group.
-Injunctive norms, on the other hand, are about what people believe they should or shouldn't do, based on the group's values or expectations for behavior.
-These norms carry a sense of approval or disapproval, guiding behavior by suggesting what is socially acceptable or frowned upon.
-Both types of norms are powerful in shaping individual actions by providing a roadmap of 'what is done' and 'what ought to be done,' helping maintain social order and cohesion by aligning personal conduct with group expectations.
+Regret, defined as an emotional response to the memory of past decisions or actions that one wishes had been different, significantly impacts our daily lives.
+This paper delves into the factors that influence the intensity of regret, with a specific focus on social norms and past behaviors.
+Understanding these elements is essential for a clearer grasp of the decision-making process and the emotions that result from it.
 
-This study delves into how people make decisions, feel about their choices, and, how they deal with regret when things don't go as planned.
-How much more would people regret their decisions just because they're not what you usually do?
-This idea stems from a concept called norm theory, which suggests that people feel extra bad about outcomes when they result from actions that are not typical for them.
-Two main experiments was introduced to unravel this complex bundle of decisions and emotions.
-The first experiment places people in a scenario where they decide whether or not to pick up a hitchhiker, testing how much one would regret the decision based on whether it's something they normally do.
-The second experiment changes the setting to driving, where choosing a different route leads to an accident, and looks at how this break from routine affects their feelings of regret.
+We categorize social norms into two main types: descriptive norms, which concern beliefs about common behaviors within a group, and injunctive norms, which relate to beliefs about what behaviors are approved or disapproved by others.
+Our research involves two experiments aimed at assessing the effect of these norms on regret.
+The first experiment looks into the role of these norms in the context of deciding whether or not to pick up a hitchhiker, while the second addresses the consequences of selecting an unconventional driving route.
 
-By picking apart these scenarios, this paper aims to shed light on the tricky relationship between stepping out of one's comfort zone in decision-making and the regret that can follow, especially when things don't turn out well.
-The goal of this research is to contribute to the broader conversation on how people's minds work in these situations, improving understanding of why people feel the way they do after making certain choices.
-This insight is not just for psychologists but for anyone interested in the little decisions people make every day and how they shape their feelings and lives.
-With all been said, we predicted that someone who often picks up hitchhikers would be seen as doing something unusual (against descriptive norms) and might be judged more harshly by others (against injunctive norms).
-We also introduced a new aspect to the original experiment, looking at how these social norms and a person's usual actions stir up feelings, both good and bad.
+The findings from these experiments reveal that unusual or unexpected events tend to provoke stronger feelings of regret than routine behaviors.
+However, the influence of both injunctive and descriptive social norms on the level of regret was found to be less significant.
+These results highlight how deviations from usual behavior, particularly when in conflict with social norms, can intensify feelings of regret.
+This insight is crucial for a better understanding of the cognitive processes involved in evaluating decisions and the emotional aftermath.
 
-The remainder of this paper is structured as follows.
-
-@sec-data
-@sec-res
-@sec-dis
+The organization of this paper is methodical and clear.
+We start with an overview of the theoretical framework concerning social norms and their connection to regret in the introduction.
+The methodology and findings from our experiments are detailed in the subsequent data section.
+The discussion section then addresses the implications of our results, outlines the limitations of the study, and suggests directions for future research.
 

### 2024-02-14 outputs/paper/paper.qmd
references.bib updated, intro updated

@@
 
-Regret, defined as an emotional response to the memory of past decisions or actions that one wishes had been different, significantly impacts our daily lives.
+Regret, defined as an emotional response to the memory of past decisions or actions that one wishes had been different, significantly impacts our daily lives [@citeRegret].
 This paper delves into the factors that influence the intensity of regret, with a specific focus on social norms and past behaviors.
@@
 
-We categorize social norms into two main types: descriptive norms, which concern beliefs about common behaviors within a group, and injunctive norms, which relate to beliefs about what behaviors are approved or disapproved by others.
+We categorize social norms into two main types: descriptive norms, which concern beliefs about common behaviors within a group, and injunctive norms, which relate to beliefs about what behaviors are approved or disapproved by others [@citeUNICEF].
 Our research involves two experiments aimed at assessing the effect of these norms on regret.
@@
 Data used in this paper was cleaned and processed with the programming language R [@citeR].
-Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `dplyr` [@citeDplyr], `readr` [@citeReadr], `knitr` [@citeknitr].
+Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `dplyr` [@citeDplyr], `readr` [@citeReadr], `knitr` [@citeknitr], 'kableExtra' [@citeKableExtra], 'reshape2' [@citeReshape2].
 

### 2024-02-14 outputs/paper/paper.qmd
source and method section are written

@@
 
-The data used in this paper was a replication of experimental data presented in the @citeOriginalPaper paper.
+The data utilized in this study replicates experimental findings from the research documented in @citeOriginalPaper.
+This original research itself is a replication of three seminal experiments designed to explore the influence of past behaviors on regret.
+The initial two experiments were pioneered by Kahneman and Miller in 1986, with the third experiment conducted by Miller and McFarland in the same year.
+Furthermore, the original study expands upon these foundational experiments.
+For the first experiment, it introduces an additional parameter to measure the overall emotional reaction of participants, considering the impact of both types of social norms.
+In the second experiment, a "Luck" measure was added to determine if the way people estimate probabilities influences their sense of what is typical.
+The expansion of the third experiment is not covered in this paper, as our examination is limited to the initial two experiments.
 
 ## Method
+In the first experiment, participants were presented with a well-known scenario by Kahneman and Miller, often referred to as the hitchhiker scenario: "Mr.
+Jones almost never takes hitch-hikers in his car.
+Yesterday he gave a man a ride and was robbed.
+Mr.
+Smith frequently takes hitch-hikers in his car.
+Yesterday he gave a man a ride and was robbed" (@citeOriginalPaper).
+The original study included three comprehension questions to confirm participants' understanding of the scenario, which we will not detail here.
+To assess the influence of injunctive and descriptive social norms, participants were asked: "Whose behavior do you think is more common in society?
+Mr.Jones(1)/Mr.Smith(2)" and "Whose behavior do you think will be more criticized by others in society?
+Mr.Jones(1)/Mr.Smith(2)" respectively.
+And the level of regret is also assessed by asking the question "Who do you expect will experience greater regret over the episode?
+Mr.Jones(1)/Mr.Smith(2)" (@citeOriginalPaper).
+The additional measure of negative effect is measured by asking the question "Contemplating your previous answers about this scenario and factoring in both Mr.Jones and Mr.Smith personal routines and your perceptions of social norms and possible social criticism, who do you think overall experienced more negative feelings about the decision to take a hitch-hiker that day?
+Mr.Jones(1)/Mr.Smith(2)".
 
-### Experiment #1
+The second experiment involved another scenario by Kahneman and Miller, known as the car accident scenario: "Mr.
+Adams was involved in an accident when driving home after work on his regular route.
+Mr.
+White was involved in a similar accident when driving on a route that he only takes when he wants a change of scenery" (@citeOriginalPaper).
+As with the first experiment, comprehension questions were included but are not discussed here.
+The feeling of regret is assessed by asking the question "Who is more upset over the accident?
+Mr.Jones(1)/Mr.Smith(2)".
+The "Luck" measure is examined by the question "Which of the two do you think is less lucky?
+Mr.Jones(1)/Mr.Smith(2)".
+
+# Results
+## Experiment 1
 
@@
 
-### Experiment #2
+## Experiment 2
 
@@
 
-# Results {#sec-res}
-
-## Experiment #1
-
-## Experiment #2
-
 # Discussion {#sec-dis}
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-## Diagnostics
-
 \newpage

### 2024-02-14 outputs/paper/paper.qmd
example table added

@@
 The feeling of regret is assessed by asking the question "Who is more upset over the accident?
-Mr.Jones(1)/Mr.Smith(2)".
+Mr.Adams(1)/Mr.White(2)".
 The "Luck" measure is examined by the question "Which of the two do you think is less lucky?
-Mr.Jones(1)/Mr.Smith(2)".
+Mr.Adams(1)/Mr.White(2)".
+
+## Attibutes
+The survey dataset initially contained 52 variables.
+After excluding variables related to the third experiment and comprehension questions, the dataset was narrowed down to 6 key variables for analysis.
+For the first experiment, 'Sc1_regret' represents participant responses to the regret question.
+'Sc1_socnorms1' and 'Sc1_socnorms2' correspond to the assessment of injunctive and descriptive social norms, respectively.
+'Sc1_combinednorms' is an additional variable introduced to measure the overall negative emotional response.
+As mentioned above, 1 in the dataset represents the Mr.Jones, where 2 represents Mr.Smith.
+
+In relation to the second experiment, 'Sc2_regret' denotes the responses to the regret question.
+The 'Sc2_lucky' variable is an added measure corresponding to participants' perceptions of luck as introduced in the "Luck" question.
+And 1 stands for Mr.Adams, where 2 is Mr.White.
+
+The top ten rows of filtered data with variables renamed is shown in @tbl-filtereddata.
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: tbl-filtereddata
+#| tbl-cap: top 10 rows of filtered data
+
+example_data <- data %>%
+  select(Sc1_regret, sc1_socnorms1, sc1_socnorms2, sc1_combinednorms, Sc2_regret, Sc2_lucky) %>%
+  rename(
+    "Experiment 1 regret" = Sc1_regret,
+    "Injuctive norms" = sc1_socnorms1,
+    "Descriptive norms" = sc1_socnorms2,
+    "Negative effect" = sc1_combinednorms,
+    "Experiment 2 regret" = Sc2_regret,
+    "Luck" = Sc2_lucky
+  ) %>%
+  head(10)  # Selecting only the top 10 rows for the table
+
+# Generating the table using kable and kableExtra for styling
+example_kable <- kable(example_data, format = "latex", booktabs = TRUE, align='c') %>%
+  kable_styling(full_width = FALSE, font_size = 7) 
+example_kable
+```
 
@@
 
-## Attibutes
-
-### Experiment #1
-
-### Experiment #2
-
-# Discussion {#sec-dis}
+# Discussion
 

### 2024-02-14 outputs/paper/paper.qmd
result section updated

@@
 
-# Results
+# Results {#sec-res}
+=======
 ## Experiment 1
+
+### Regret
+
+From what we learned before, we expected that participants would believe Mr.
+Jones, who rarely picks up hitchhikers, would feel more regret than Mr.
+Smith, who does so often.
+It turns out, most participants did think Mr.
+Jones would feel worse (315 responses, or 92%) compared to Mr.
+Smith (27 responses, or 8%), as shown in the @fig-experiment1bar.
+
+### Social norms
+
+Taking our research a step further, we hypothesized that participants would see Mr.
+Jones's behavior (not picking up hitchhikers) as more typical or normal, and Mr.
+Smith's behavior (picking up hitchhikers) as less typical and more likely to be judged by others.
+This was indeed what we found: most people thought Mr.
+Jones's actions were more common (326 responses, or 95%) and Mr.
+Smith's actions were more likely to face social criticism (310 responses, or 91%), as shown in the tbl-experiment1.
 
@@
 ## Experiment 2
+
+### Regret
+
+We had a hunch that participants would feel Mr.
+White, who had an accident on an unusual route home, would regret his decision more than Mr.
+Adams, who had his accident on his regular drive home.
+And it turns out, that's exactly what most participants thought: 277 of them (81%) associated a higher level of regret with Mr.
+White's unusual decision, while only 65 of them (19%) thought Mr.
+Adams would feel more regret, as shown in the @fig-experiment2bar.
+
+Thus, our study confirmed our guess: Mr.
+White's one-off decision led to more regret among our participants than Mr.
+Adams's everyday routine, suggesting that when bad things happen on a path less traveled, they hit us harder.
 

### 2024-02-14 outputs/paper/paper.qmd
result for experiment 1 added

--- 2024-02-14 outputs/paper/paper.qmd
@@
 # Generating the table using kable and kableExtra for styling
-example_kable <- kable(example_data, format = "latex", booktabs = TRUE, align='c') %>%
+example_kable <- kable(example_data, format = "latex", booktabs = TRUE, align='c') %>% # align the text in the middle
   kable_styling(full_width = FALSE, font_size = 7) 
@@
 ```
+
+@tbl-experiment1 and @fig-experiment1bar display results from experiment 1.
+Mr.
+Jones, who performs an action that is an exception to his routine, reports a higher regret level of 92.1%.
+Most participants rated Mr.Jones' behavior to be more common (descriptive) as he never gave ride to hitchhiker (95.3) and therefore to be less criticized (injunctive) by the society (9.4%).
+Mr.Jones is thought to experience more negative feelings about the decision made that day combining previous answers to the two social norms (92.7%).
+Mr.Smith, associated with routine behavior, shows markedly lower levels in these aspects, participants rated 7.9%, 4.7%, 90.6% and 7.3% for regret, descriptive social norm, injunctive social norm, and negative effect respectively.
+Indicating that although Mr.Smith is thought to be criticized more in the society, he would experience less regret and negative effect overall.
+
+In academic terms, the disparity in responses suggests that individuals who engage in behaviors that are not typical for them (Mr.
+Jones) are more likely to experience and be perceived as having stronger regret and more negative emotional responses.
+The consistency in this pattern across different measures indicates a robust finding, aligning with psychological theories that predict increased emotional consequences for actions that deviate from one's own or society's expectations.
+The bar plot @fig-experiment1bar quantitatively illustrates these findings.
+Mr.
+Jones's experiences with regret and negative affect are visually represented by significantly taller bars compared to Mr.
+Smith's, reflecting the numerical data.
 

### 2024-02-14 outputs/paper/paper.qmd

@@
 
+### Regret
+
+We had a hunch that participants would feel Mr.
+White, who had an accident on an unusual route home, would regret his decision more than Mr.
+Adams, who had his accident on his regular drive home.
+And it turns out, that's exactly what most participants thought: 277 of them (81%) associated a higher level of regret with Mr.
+White's unusual decision, while only 65 of them (19%) thought Mr.
+Adams would feel more regret, as shown in the @fig-experiment2bar.
+
+Thus, our study confirmed our guess: Mr.
+White's one-off decision led to more regret among our participants than Mr.
+Adams's everyday routine, suggesting that when bad things happen on a path less traveled, they hit us harder.
+
 ```{r}

### 2024-02-15 outputs/paper/paper.qmd
result added
results of the two experiments added, discussion section structure completed

--- 2024-02-14 outputs/paper/paper.qmd
@@
 
-In academic terms, the disparity in responses suggests that individuals who engage in behaviors that are not typical for them (Mr.
+The disparity in responses suggests that individuals who engage in behaviors that are not typical for them (Mr.
 Jones) are more likely to experience and be perceived as having stronger regret and more negative emotional responses.
@@
 
+@tbl-experiment2 and @fig-experiment2bar illustrate the outcomes of experiment 2, which addresses perceptions of regret and luck in decision-making under varying conditions of routine and exception.
+
+Mr.
+Adams, who represents routine behavior, exhibits a regret level of 18.8%, suggesting that participants view routine actions as less regrettable.
+His luck perception stands at 32.9%, indicating that participants may see routine choices as less influenced by luck.
+On the contrary, Mr.
+White, who represents an exception to routine, is associated with a markedly higher regret level at 81.2%.
+This suggests that non-routine actions are perceived as significantly more regrettable.
+Furthermore, his luck perception is at 67.1%, which could reflect a belief that deviations from routine are more susceptible to chance and less within an individual's control.
+
+@fig-experiment2bar graphically accentuates these findings, with much higher height bars for Mr.White in terms of regret and luck.
+This visual comparison conveys the statistical differences and underscores the paper's exploration of the psychological impacts of routine versus exceptional decisions.
+This visualization complements the statistical data by providing an immediate, comparative view of the emotional weight and perceived randomness assigned to each individual's experience.
+
 # Discussion
 
-## First discussion point
+## Replication and estimand
 
-## Second discussion point
+## Measurement validity
 
-## Third discussion point
+## Ethical Considerations and Bias
 
-## Weaknesses and next steps
+## Limitations and Future Research
+
+## Conclusion
 

### 2024-02-15 outputs/paper/paper.qmd
discussion section completed

@@
 ## Replication and estimand
+In replicating the classic experiments by Kahneman and Miller (1986), our study sought to precisely estimate the effect of deviations from routine on the experience of regret — a clearly defined estimand.
+This estimand was articulated in the introduction and guided the experimental design to ensure the validity of our findings.
 
 ## Measurement validity
+The validity of our measurements was paramount, as we endeavored to capture quantifiable aspects of regret that directly relate to our estimand.
+Our measures of regret, social norms, and the added dimension of luck were developed to reflect the actual constructs under study, adhering to the principles outlined by Flake and Fried (2020).
+This rigorous approach to measurement ensures that the quantified variables are not only appropriate but also reliably represent the concepts they are intended to measure.
 
 ## Ethical Considerations and Bias
+Ethical considerations in replicating behavioral experiments are critical.
+We ensured the ethical treatment of participants and transparency in our methodology, especially considering the subjective nature of self-reported data on regret.
+The study's design attempted to minimize bias, but we acknowledge that no research is free from all bias, including potential measurement errors or the influence of the study's context on participants' responses.
 
 ## Limitations and Future Research
+A limitation of the current study includes the use of vignettes, which may not fully capture the complexity of real-life decisions and their emotional consequences.
+Future studies might explore the phenomenon in naturalistic settings or employ longitudinal designs to observe the persistence of regret over time.
+Additional research could also examine whether different types of regret have varying impacts on future decision-making and behavior change.
 
 ## Conclusion
+Our study contributes to the body of literature on decision-making and emotion by providing a refined estimation of the exceptionality effect's impact on regret.
+The robustness of the effect across replications reinforces its relevance in psychological research.
+Moving forward, there is a need for more nuanced investigations that consider the complexity of human behavior and the myriad factors that shape our emotional responses to our decisions.
 

### 2024-02-15 outputs/paper/paper.qmd
scripts added

@@
 
+# calculate each percentage
 smith_regret_p <- round((mr_smith_counts / total_counts_1) * 100, 1)

### 2024-02-15 replication/scripts/fig1_replication.R

@@ -1,5 +1,18 @@
-# experiment1bar
+# load packages
+library(tidyverse)
+library(ggplot2)
+library(janitor)
+library(dplyr)
+library(readr)
+library(knitr)
+library(kableExtra)
+library(reshape2)
 
+# read in data
+file_path <- "inputs/data/raw_data.csv"
+data <- read_csv(file_path, show_col_types = FALSE)
+
+# calculate each percentage
 smith_regret_p <- round((mr_smith_counts / total_counts_1) * 100, 1)
 smith_injunctive_p <- round((sum(data_filtered_1$sc1_socnorms1 == 2) / total_counts_1) * 100, 1)
 smith_descriptive_p <- round((sum(data_filtered_1$sc1_socnorms2 == 2) / total_counts_1) * 100, 1)
@@ -30,39 +43,6 @@ ggplot(data_long, aes(x = Category, y = Percentage, fill = Condition)) +
   scale_fill_manual(values = c("Smith_Routine" = "grey", "Jones_Exception" = "#505050"), 
                     labels = c("Mr.Smith(Routine)", "Mr.Jones(Exception)")) +
   scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20)) +
-  theme_minimal() +
-  labs(x = "", y = "Percentage", fill = "Condition") +
-  theme(axis.text.x = element_text(hjust = 1, vjust = 1, colour = "black",size = 9), 
-        legend.position = "top", legend.title = element_blank())
-
-
-
-# experiment2bar
-
-adams_regret_p <- round((mr_adams_counts / total_counts_2) * 100, 1)
-adams_luck_p <- round((sum(data_filtered_2$Sc2_lucky == 1) / total_counts_2) * 100, 1)
-white_regret_p <- round((sum(data_filtered_2$Sc2_regret == 2) / total_counts_2) * 100, 1)
-white_luck_p <- round((sum(data_filtered_2$Sc2_lucky == 2) / total_counts_2) * 100, 1)
-
-# Create a data frame with the provided data
-data <- data.frame(
-  Category = c("Regret", "Luck"),
-  Adams_Routine = c(adams_regret_p, adams_luck_p),
-  White_Exception = c(white_regret_p, white_luck_p)
-)
-
-# Convert the data from wide to long format
-data_long <- melt(data, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")
-
-# Reorder the factor levels of Category so that 'Regret' comes first
[…]

### 2024-02-15 replication/scripts/fig2_replication.R

@@ -1,0 +1,43 @@
+# load packages
+library(tidyverse)
+library(ggplot2)
+library(janitor)
+library(dplyr)
+library(readr)
+library(knitr)
+library(kableExtra)
+library(reshape2)
+
+# read in data
+file_path <- "inputs/data/raw_data.csv"
+data <- read_csv(file_path, show_col_types = FALSE)
+
+# calculate each percentage
+adams_regret_p <- round((mr_adams_counts / total_counts_2) * 100, 1)
+adams_luck_p <- round((sum(data_filtered_2$Sc2_lucky == 1) / total_counts_2) * 100, 1)
+white_regret_p <- round((sum(data_filtered_2$Sc2_regret == 2) / total_counts_2) * 100, 1)
+white_luck_p <- round((sum(data_filtered_2$Sc2_lucky == 2) / total_counts_2) * 100, 1)
+
+# Create a data frame with the provided data
+data <- data.frame(
+  Category = c("Regret", "Luck"),
+  Adams_Routine = c(adams_regret_p, adams_luck_p),
+  White_Exception = c(white_regret_p, white_luck_p)
+)
+
+# Convert the data from wide to long format
+data_long <- melt(data, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")
+
+# Reorder the factor levels of Category so that 'Regret' comes first
+data_long$Category <- factor(data_long$Category, levels = c("Regret", "Luck"))
+
+# Create the bar plot with the categories reordered
+ggplot(data_long, aes(x = Category, y = Percentage, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.7) +
+  scale_fill_manual(values = c("Adams_Routine" = "grey", "White_Exception" = "#505050"), 
+                    labels = c("Mr.Adams(Routine)", "Mr.White(Exception)")) +
+  scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20)) + # label the y-axis every 20%
+  theme_minimal() +
+  labs(x = "", y = "Percentage", fill = "Condition") +
+  theme(axis.text.x = element_text(hjust = 1, vjust = 1, colour = "black",size = 9), 
+        legend.position = "top", legend.title = element_blank())

### 2024-02-15 replication/scripts/table1_replication.R

@@ -1,0 +1,49 @@
+# load packages
+library(tidyverse)
+library(ggplot2)
+library(janitor)
+library(dplyr)
+library(readr)
+library(knitr)
+library(kableExtra)
+library(reshape2)
+
+# read in data
+file_path <- "inputs/data/raw_data.csv"
+data <- read_csv(file_path, show_col_types = FALSE)
+
+# Filter the data to drop NA values
+data_filtered_1 <- data %>%
+  filter(!is.na(Sc1_regret) & !is.na(sc1_socnorms1) & !is.na(sc1_socnorms2) & !is.na(sc1_combinednorms))
+
+# Calculate the total counts and counts for Mr.Jones (1) and Mr.Smith (2)
+total_counts_1 <- nrow(data_filtered_1)
+mr_jones_counts <- sum(data_filtered_1$Sc1_regret == 1)
+mr_smith_counts <- total_counts_1 - mr_jones_counts
+
+# Calculate the counts and percentages for Mr.Jones
+mr_jones_data <- data.frame(
+  Group = "Mr.Jones(exception)",
+  Regret = paste(mr_jones_counts, "(", round((mr_jones_counts / total_counts_1) * 100, 1), "%)", sep = ""),
+  Injunctive = paste(sum(data_filtered_1$sc1_socnorms1 == 1), "(", round((sum(data_filtered_1$sc1_socnorms1 == 1) / total_counts_1) * 100, 1), "%)", sep = ""),
+  Descriptive = paste(sum(data_filtered_1$sc1_socnorms2 == 1), "(", round((sum(data_filtered_1$sc1_socnorms2 == 1) / total_counts_1) * 100, 1), "%)", sep = ""),
+  `Negative Effect` = paste(sum(data_filtered_1$sc1_combinednorms == 1), "(", round((sum(data_filtered_1$sc1_combinednorms == 1) / total_counts_1) * 100, 1), "%)", sep = "")
+)
+
+# Calculate the counts and percentages for Mr.Smith
+mr_smith_data <- data.frame(
+  Group = "Mr.Smith(routine)",
+  Regret = paste(mr_smith_counts, "(", round((mr_smith_counts / total_counts_1) * 100, 1), "%)", sep = ""),
+  Injunctive = paste(sum(data_filtered_1$sc1_socnorms1 == 2), "(", round((sum(data_filtered_1$sc1_socnorms1 == 2) / total_counts_1) * 100, 1), "%)", sep = ""),
[…]

### 2024-02-15 replication/scripts/table2 _replication.R

@@ -1,0 +1,45 @@
+# load packages
+library(tidyverse)
+library(ggplot2)
+library(janitor)
+library(dplyr)
+library(readr)
+library(knitr)
+library(kableExtra)
+library(reshape2)
+
+# read in data
+file_path <- "inputs/data/raw_data.csv"
+data <- read_csv(file_path, show_col_types = FALSE)
+
+# Filter the data to drop NA values
+data_filtered_2 <- data %>%
+  filter(!is.na(Sc2_regret) & !is.na(Sc2_random_1) & !is.na(Sc2_random_2) & !is.na(Sc2_lucky))
+
+# Calculate the total counts and counts for Mr.Adams (1) and Mr.White (2)
+total_counts_2 <- nrow(data_filtered_2)
+mr_adams_counts <- sum(data_filtered_2$Sc2_regret == 1)
+mr_white_counts <- total_counts_2 - mr_adams_counts
+
+# Calculate the counts and percentages for Mr.Adams
+mr_adams_data <- data.frame(
+  Group = "Mr.Adams(routine)",
+  Regret = paste(mr_adams_counts, "(", round((mr_adams_counts / total_counts_2) * 100, 1), "%)", sep = ""),
+  Luck = paste(sum(data_filtered_2$Sc2_lucky == 1), "(", round((sum(data_filtered_2$Sc2_lucky == 1) / total_counts_2) * 100, 1), "%)", sep = "")
+)
+
+# Calculate the counts and percentages for Mr. White
+mr_white_data <- data.frame(
+  Group = "Mr.White(exception)",
+  Regret = paste(mr_white_counts, "(", round((mr_white_counts / total_counts_2) * 100, 1), "%)", sep = ""),
+  Luck = paste(sum(data_filtered_2$Sc2_lucky == 2), "(", round((sum(data_filtered_2$Sc2_lucky == 2) / total_counts_2) * 100, 1), "%)", sep = "")
+)
+
+# Combine the data for Mr. Jones and Mr. Smith into one dataframe
+combined_data <- rbind(mr_adams_data, mr_white_data)
+
+# Now use kable to create the table with the combined data
+kable_output <- kable(combined_data, "latex", booktabs = TRUE) %>%
+  kable_styling(full_width = FALSE, font_size = 7)
+
+kable_output

### 2024-02-15 replication/scripts/tables.R

@@ -1,94 +1,0 @@
-# filtereddata
-
-example_data <- data %>%
-  select(Sc1_regret, sc1_socnorms1, sc1_socnorms2, sc1_combinednorms, Sc2_regret, Sc2_lucky) %>%
-  rename(
-    "Experiment 1 regret" = Sc1_regret,
-    "Injuctive norms" = sc1_socnorms1,
-    "Descriptive norms" = sc1_socnorms2,
-    "Negative effect" = sc1_combinednorms,
-    "Experiment 2 regret" = Sc2_regret,
-    "Luck" = Sc2_lucky
-  ) %>%
-  head(10)  # Selecting only the top 10 rows for the table
-
-# Generating the table using kable and kableExtra for styling
-example_kable <- kable(example_data, format = "latex", booktabs = TRUE, align='c') %>% # align the text in the middle
-  kable_styling(full_width = FALSE, font_size = 7) 
-example_kable
-
-
-
-# experiment1
-
-# Filter the data to drop NA values
-data_filtered_1 <- data %>%
-  filter(!is.na(Sc1_regret) & !is.na(sc1_socnorms1) & !is.na(sc1_socnorms2) & !is.na(sc1_combinednorms))
-
-# Calculate the total counts and counts for Mr.Jones (1) and Mr.Smith (2)
-total_counts_1 <- nrow(data_filtered_1)
-mr_jones_counts <- sum(data_filtered_1$Sc1_regret == 1)
-mr_smith_counts <- total_counts_1 - mr_jones_counts
-
-# Calculate the counts and percentages for Mr.Jones
-mr_jones_data <- data.frame(
-  Group = "Mr.Jones(exception)",
-  Regret = paste(mr_jones_counts, "(", round((mr_jones_counts / total_counts_1) * 100, 1), "%)", sep = ""),
-  Injunctive = paste(sum(data_filtered_1$sc1_socnorms1 == 1), "(", round((sum(data_filtered_1$sc1_socnorms1 == 1) / total_counts_1) * 100, 1), "%)", sep = ""),
-  Descriptive = paste(sum(data_filtered_1$sc1_socnorms2 == 1), "(", round((sum(data_filtered_1$sc1_socnorms2 == 1) / total_counts_1) * 100, 1), "%)", sep = ""),
[…]

### 2024-02-15 outputs/paper/paper.qmd
completed

@@
   - Yihang Cai
-thanks: "Code and data are available at: https://github.com/iJustinn/Impact_of_Past_Behavior_Normality_on_Regret.git"
+thanks: "Code and data are available at: https://github.com/iJustinn/Impact_of_Past_Behavior_Normality_on_Regret.git, the original paper can be found at https://doi.org/10.1080/02699931.2018.1504747"
 date: today
-date-format: long
-abstract: "To be written after we finished the Discussion section"
+date-format: "long" 
+abstract: "The paper investigates the influence of deviation from routine versus social norms on regret.
+Through replicating two significant experiments, it found that deviations, or the exceptionality effect, significantly impact regret levels.
+Participants perceived those who diverged from routine behaviors as experiencing more regret, a finding consistent across two experiments and further underscored by measurements of negative effect and luck.
+This study enriches the understanding of decision-making and emotions, emphasizing the pivotal role of personal behavioral norms in shaping regret, thereby contributing to psychological research's knowledge base."
 format: pdf
+output: bookdown::pdf_document2
 toc: true
-toc-depth: 3
 number-sections: true
@@
 #| echo: false
+
+knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
 
@@
 
+We are trying to measure which of the two factors has a more significant effect on the level of regret (estimand): social norms and past behaviors(exceptionality effect).
 We categorize social norms into two main types: descriptive norms, which concern beliefs about common behaviors within a group, and injunctive norms, which relate to beliefs about what behaviors are approved or disapproved by others [@citeUNICEF].
@@
 ## Method
+
 In the first experiment, participants were presented with a well-known scenario by Kahneman and Miller, often referred to as the hitchhiker scenario: "Mr.
@@
 ## Attibutes
-The survey dataset initially contained 52 variables.
-After excluding variables related to the third experiment and comprehension questions, the dataset was narrowed down to 6 key variables for analysis.
-For the first experiment, 'Sc1_regret' represents participant responses to the regret question.
-'Sc1_socnorms1' and 'Sc1_socnorms2' correspond to the assessment of injunctive and descriptive social norms, respectively.
-'Sc1_combinednorms' is an additional variable introduced to measure the overall negative emotional response.
-As mentioned above, 1 in the dataset represents the Mr.Jones, where 2 represents Mr.Smith.
-
-In relation to the second experiment, 'Sc2_regret' denotes the responses to the regret question.
-The 'Sc2_lucky' variable is an added measure corresponding to participants' perceptions of luck as introduced in the "Luck" question.
-And 1 stands for Mr.Adams, where 2 is Mr.White.
-
-The top ten rows of filtered data with variables renamed is shown in @tbl-filtereddata.
 
@@
 #| warning: false
-#| fig-pos: 'h'
 #| label: tbl-filtereddata
@@
 
+The survey dataset initially contained 52 variables.
+After excluding variables related to the third experiment and comprehension questions, the dataset was narrowed down to 6 key variables for analysis.
+For the first experiment, 'Sc1_regret' represents participant responses to the regret question.
+'Sc1_socnorms1' and 'Sc1_socnorms2' correspond to the assessment of injunctive and descriptive social norms, respectively.
+'Sc1_combinednorms' is an additional variable introduced to measure the overall negative emotional response.
+As mentioned above, 1 in the dataset represents the Mr.Jones, where 2 represents Mr.Smith.
+
+In relation to the second experiment, 'Sc2_regret' denotes the responses to the regret question.
+The 'Sc2_lucky' variable is an added measure corresponding to participants' perceptions of luck as introduced in the "Luck" question.
+And 1 stands for Mr.Adams, where 2 is Mr.White.
+
+The top ten rows of filtered data with variables renamed is shown in @tbl-filtereddata.
+
+\newpage
+
 # Results
+
 ## Experiment 1
+
+@tbl-experiment1 displays results from experiment 1.
+Mr.
+Jones, who performs an action that is an exception to his routine, reports a higher regret level of 92.1%.
+Most participants rated Mr.Jones' behavior to be more common (descriptive) as he never gave ride to hitchhiker (95.3) and therefore to be less criticized (injunctive) by the society (9.4%).
+Mr.Jones is thought to experience more negative feelings about the decision made that day combining previous answers to the two social norms (92.7%).
+Mr.Smith, associated with routine behavior, shows markedly lower levels in these aspects, participants rated 7.9%, 4.7%, 90.6% and 7.3% for regret, descriptive social norm, injunctive social norm, and negative effect respectively.
+Indicating that although Mr.Smith is thought to be criticized more in the society, he would experience less regret and negative effect overall.
 
@@
 #| warning: false
-#| fig-pos: 'h'
+#| fig-pos: 'H'
 #| label: tbl-experiment1
@@
 
+The disparity in responses suggests that individuals who engage in behaviors that are not typical for them (Mr.
+Jones) are more likely to experience and be perceived as having stronger regret and more negative emotional responses.
+The consistency in this pattern across different measures indicates a robust finding, aligning with psychological theories that predict increased emotional consequences for actions that deviate from one's own or society's expectations.
+
 ```{r}
@@
 #| warning: false
-#| fig-pos: 'h'
 #| label: fig-experiment1bar
@@
 ```
-
-@tbl-experiment1 and @fig-experiment1bar display results from experiment 1.
-Mr.
-Jones, who performs an action that is an exception to his routine, reports a higher regret level of 92.1%.
-Most participants rated Mr.Jones' behavior to be more common (descriptive) as he never gave ride to hitchhiker (95.3) and therefore to be less criticized (injunctive) by the society (9.4%).
-Mr.Jones is thought to experience more negative feelings about the decision made that day combining previous answers to the two social norms (92.7%).
-Mr.Smith, associated with routine behavior, shows markedly lower levels in these aspects, participants rated 7.9%, 4.7%, 90.6% and 7.3% for regret, descriptive social norm, injunctive social norm, and negative effect respectively.
-Indicating that although Mr.Smith is thought to be criticized more in the society, he would experience less regret and negative effect overall.
-
-The disparity in responses suggests that individuals who engage in behaviors that are not typical for them (Mr.
-Jones) are more likely to experience and be perceived as having stronger regret and more negative emotional responses.
-The consistency in this pattern across different measures indicates a robust finding, aligning with psychological theories that predict increased emotional consequences for actions that deviate from one's own or society's expectations.
 The bar plot @fig-experiment1bar quantitatively illustrates these findings.
@@
 
+@tbl-experiment2 and @fig-experiment2bar illustrate the outcomes of experiment 2, which addresses perceptions of regret and luck in decision-making under varying conditions of routine and exception.
+
+Mr.
+Adams, who represents routine behavior, exhibits a regret level of 18.8%, suggesting that participants view routine actions as less regrettable.
+His luck perception stands at 32.9%, indicating that participants may see routine choices as less influenced by luck.
+On the contrary, Mr.
+White, who represents an exception to routine, is associated with a markedly higher regret level at 81.2%.
+This suggests that non-routine actions are perceived as significantly more regrettable.
+Furthermore, his luck perception is at 67.1%, which could reflect a belief that deviations from routine are more susceptible to chance and less within an individual's control.
+
 ```{r}
 #| message: false
+#| fig-pos: 'h'
 #| echo: false
 #| warning: false
-#| fig-pos: 'h'
 #| label: tbl-experiment2
@@
 
+@fig-experiment2bar graphically accentuates these findings, with much higher height bars for Mr.White in terms of regret and luck.
+This visual comparison conveys the statistical differences and underscores the paper's exploration of the psychological impacts of routine versus exceptional decisions.
+This visualization complements the statistical data by providing an immediate, comparative view of the emotional weight and perceived randomness assigned to each individual's experience.
+
 ```{r}
@@
 #| warning: false
-#| fig-pos: 'h'
 #| label: fig-experiment2bar
@@
 
+# calculate each percentage
 adams_regret_p <- round((mr_adams_counts / total_counts_2) * 100, 1)
@@
 
-@tbl-experiment2 and @fig-experiment2bar illustrate the outcomes of experiment 2, which addresses perceptions of regret and luck in decision-making under varying conditions of routine and exception.
-
-Mr.
-Adams, who represents routine behavior, exhibits a regret level of 18.8%, suggesting that participants view routine actions as less regrettable.
-His luck perception stands at 32.9%, indicating that participants may see routine choices as less influenced by luck.
-On the contrary, Mr.
-White, who represents an exception to routine, is associated with a markedly higher regret level at 81.2%.
-This suggests that non-routine actions are perceived as significantly more regrettable.
-Furthermore, his luck perception is at 67.1%, which could reflect a belief that deviations from routine are more susceptible to chance and less within an individual's control.
-
-@fig-experiment2bar graphically accentuates these findings, with much higher height bars for Mr.White in terms of regret and luck.
-This visual comparison conveys the statistical differences and underscores the paper's exploration of the psychological impacts of routine versus exceptional decisions.
-This visualization complements the statistical data by providing an immediate, comparative view of the emotional weight and perceived randomness assigned to each individual's experience.
-
 # Discussion
@@
 ## Replication and estimand
-In replicating the classic experiments by Kahneman and Miller (1986), our study sought to precisely estimate the effect of deviations from routine on the experience of regret — a clearly defined estimand.
+
+In replicating the classic experiments by Kahneman and Miller (1986), our study sought to estimate the effect of social norms and past behaviors on the experience of regret (which of the two is more important) — a clearly defined estimand.
 This estimand was articulated in the introduction and guided the experimental design to ensure the validity of our findings.
@@
 ## Measurement validity
-The validity of our measurements was paramount, as we endeavored to capture quantifiable aspects of regret that directly relate to our estimand.
-Our measures of regret, social norms, and the added dimension of luck were developed to reflect the actual constructs under study, adhering to the principles outlined by Flake and Fried (2020).
+
+The validity of our measurements was crucial, as we endeavored to capture quantifiable aspects of regret that directly relate to our estimand.
+Our measures of regret, social norms, and the added dimension of luck and negative effect were developed to reflect the actual constructs under study.
 This rigorous approach to measurement ensures that the quantified variables are not only appropriate but also reliably represent the concepts they are intended to measure.
@@
 ## Ethical Considerations and Bias
+
 Ethical considerations in replicating behavioral experiments are critical.
-We ensured the ethical treatment of participants and transparency in our methodology, especially considering the subjective nature of self-reported data on regret.
+The original paper recruited American Amazon Mechanical Turk (MTurk) participants online, and ensured that the participants took the survey in the first experiment could not take in the second experiment.
+Thus, eliminating any possible effects between the two experiments.
 The study's design attempted to minimize bias, but we acknowledge that no research is free from all bias, including potential measurement errors or the influence of the study's context on participants' responses.
@@
 ## Limitations and Future Research
+
 A limitation of the current study includes the use of vignettes, which may not fully capture the complexity of real-life decisions and their emotional consequences.
@@
 
-## Conclusion
-Our study contributes to the body of literature on decision-making and emotion by providing a refined estimation of the exceptionality effect's impact on regret.
+## Result and Conclusion
+
+The study's findings suggest that deviations from routine have a more pronounced impact on regret than social norms.
+Both experiments showed that participants believed individuals who diverged from their usual behavior (Mr.
+Jones and Mr.
+White) felt more regret compared to those who did not.
+This was further supported by the negative effect measurement, which took into account both injunctive and descriptive social norms.
+Additionally, participants considered Mr.
+White to be less fortunate than Mr.
+Adams, suggesting that routine actions might lead to fewer regrets.
+This research adds valuable insight into how exceptional actions influence regret, reinforcing the significance of the exceptionality effect in understanding emotional responses in decision-making.
 The robustness of the effect across replications reinforces its relevance in psychological research.
-Moving forward, there is a need for more nuanced investigations that consider the complexity of human behavior and the myriad factors that shape our emotional responses to our decisions.
 

### 2024-02-15 replication/scripts/simulation.R
simulation file added

@@ -1,0 +1,51 @@
+library(tidyverse)
+library(ggplot2)
+library(janitor)
+library(dplyr)
+library(readr)
+library(knitr)
+library(kableExtra)
+library(reshape2)
+
+# experiment 1 simulation table
+data <- data.frame(
+  Category = c("Regret", "Negative_Affect"),
+  Smith_Routine = c(10, 8),
+  Jones_Exception = c(90, 92)
+)
+
+kable(data, caption = "Descriptive Statistics", align = c('l', 'c', 'c'))
+
+# experiment 1 simulation bar plot
+
+# Convert the data from wide to long format
+data_long <- melt(data, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")
+
+# Create the bar plot
+ggplot(data_long, aes(x = Category, y = Percentage, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
+  scale_fill_manual(values = c("Smith_Routine" = "blue", "Jones_Exception" = "red")) +
+  theme_minimal() +
+  labs(x = "", y = "Percentage", fill = "Condition") +
+  theme(legend.position = "bottom")
+
+# experiment 2 simulation table
+data_2 <- data.frame(
+  Category = c("Regret", "Luck"),
+  Adams_Routine = c(12, 30),
+  White_Exception = c(88, 70)
+)
+
+kable(data_2, caption = "Descriptive Statistics", align = c('l', 'c', 'c'))
+
+# experiment 2 simulation bar plot
+# Convert the data from wide to long format
+data_long <- melt(data_2, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")
+
+# Create the bar plot
+ggplot(data_long, aes(x = Category, y = Percentage, fill = Condition)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
+  scale_fill_manual(values = c("Adams_Routine" = "blue", "White_Exception" = "red")) +
+  theme_minimal() +
+  labs(x = "", y = "Percentage", fill = "Condition") +
+  theme(legend.position = "bottom")

### 2024-02-15 replication/scripts/simulation.R
testing file changed

@@ -9,43 +9,110 @@ library(reshape2)
 
 # experiment 1 simulation table
 data <- data.frame(
-  Category = c("Regret", "Negative_Affect"),
-  Smith_Routine = c(10, 8),
-  Jones_Exception = c(90, 92)
+  Category = c("Mr.Smith", "Mr.Jones"),
+  Regret = c(10, 90),
+  Luck = c(8, 92)
 )
 
 kable(data, caption = "Descriptive Statistics", align = c('l', 'c', 'c'))
 
 # experiment 1 simulation bar plot
 
-# Convert the data from wide to long format
-data_long <- melt(data, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")
+# Convert data from wide to long format for ggplot2
+data_long <- reshape2::melt(data, id.vars = "Category", variable.name = "Outcome", value.name = "Percentage")
 
 # Create the bar plot
-ggplot(data_long, aes(x = Category, y = Percentage, fill = Condition)) +
-  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
-  scale_fill_manual(values = c("Smith_Routine" = "blue", "Jones_Exception" = "red")) +
+ggplot(data_long, aes(x = Category, y = Percentage, fill = Outcome)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Pastel1") +
   theme_minimal() +
-  labs(x = "", y = "Percentage", fill = "Condition") +
+  labs(x = "Group", y = "Percentage", fill = "Outcome") +
   theme(legend.position = "bottom")
 
 # experiment 2 simulation table
 data_2 <- data.frame(
-  Category = c("Regret", "Luck"),
-  Adams_Routine = c(12, 30),
-  White_Exception = c(88, 70)
+  Category = c("Mr.Adams", "Mr.White"),
+  Regret = c(12, 88),
+  Luck = c(30, 70)
 )
 
 kable(data_2, caption = "Descriptive Statistics", align = c('l', 'c', 'c'))
 
 # experiment 2 simulation bar plot
-# Convert the data from wide to long format
-data_long <- melt(data_2, id.vars = "Category", variable.name = "Condition", value.name = "Percentage")
+# Convert data from wide to long format for ggplot2
+data_long <- reshape2::melt(data_2, id.vars = "Category", variable.name = "Outcome", value.name = "Percentage")
 
 # Create the bar plot
[…]

### 2024-02-15 replication/scripts/testing.R

@@ -1,28 +1,38 @@
 # Load the dataset into R
-file_path <- "../../inputs/data/raw_data.csv"
-my_data <- read_csv(file_path, show_col_types = FALSE)
+file_path <- "inputs/data/raw_data.csv"
+data <- read_csv(file_path, show_col_types = FALSE)
 
-# Check if 'age' contains only integer values
-is_age_integer <- all(my_data$age == floor(my_data$age))
+# Check if selected columns contain only the values 1 and 2
+selected_data <- data %>%
+  select(Sc1_regret, sc1_socnorms1, sc1_socnorms2, sc1_combinednorms, Sc2_regret, Sc2_lucky) %>%
+  rename(
+    "Experiment 1 regret" = Sc1_regret,
+    "Injuctive norms" = sc1_socnorms1,
+    "Descriptive norms" = sc1_socnorms2,
+    "Negative effect" = sc1_combinednorms,
+    "Experiment 2 regret" = Sc2_regret,
+    "Luck" = Sc2_lucky
+  )
 
-# Check if 'gender' contains only the values 1 and 2
-valid_genders <- unique(my_data$gender)
-is_gender_valid <- all(valid_genders %in% c(1, 2))
+selected_data <- na.omit(selected_data)
 
-# Output the results
-print(paste("Age column contains only integers:", is_age_integer))
-print(paste("Gender column contains only 1 and 2:", is_gender_valid))
+columns_to_check <- c("Experiment 1 regret", "Injuctive norms", "Descriptive norms", "Negative effect", "Experiment 2 regret", "Luck")
 
-# Optionally, identify rows with problematic 'age' data
-if(!is_age_integer) {
-  problematic_age <- my_data[my_data$age != floor(my_data$age), ]
-  print("Rows with problematic 'age' data:")
-  print(problematic_age)
-}
+results <- sapply(selected_data[columns_to_check], function(x) all(x %in% c(1, 2)))
+print(results)
 
-# Optionally, identify rows with problematic 'gender' data
-if(!is_gender_valid) {
-  problematic_gender <- my_data[!my_data$gender %in% c(1, 2), ]
-  print("Rows with problematic 'gender' data:")
-  print(problematic_gender)
-}
+# Check if selected columns contains only numeric values
+# Define the columns to check
+columns_to_check <- c("Experiment 1 regret", "Injuctive norms", "Descriptive norms", "Negative effect", "Experiment 2 regret", "Luck")
+
[…]
